Exact Symbolic Dynamic Programming for Continuous POMDPs

### 2012-10-07 cpomdp/camdp-xadd/doc/draft6-after first review3.tex
Moved NIPS 2012 latex source

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\incmargin{.5em}
\linesnumbered
\begin{algorithm}[t!]
\vspace{-.5mm}
\dontprintsemicolon
\SetKwFunction{backup}{Backup}
\SetKwFunction{genObs}{GenRelObs}
\SetKwFunction{prune}{Prune}
\SetKwFunction{remapWithPrimes}{Prime}
\Begin
{
   $V^0:=0, h:=0$\;
   \While{$h < H$}
   {
       $h:=h+1, \Gamma^h :=\emptyset$\;
       \ForEach {$a \in A$}
       {
			$\Gamma_{a}^h :=\emptyset$ \;       		
       		$\textsc{O}^{h-1} \,:=\,$ \genObs{$\Gamma^{h-1},a$}\;
       		 \ForEach {$o_i \in \textsc{O}^{h-1}$}
       		 {
				\ForEach {$\alpha_j^{h-1} \in \Gamma^{h-1}$}
       			{
   	 		  		$\alpha_j^{h'}=$ \remapWithPrimes{$\alpha_j^{h}$} \;
   	 		  		\emph{// $\forall$ $d_i \to d_i'$ and $\forall x_i \to x_i'$} \; 
   	 		    	$\Gamma_{a,o_i,j}^h \,:=\, o_i \otimes$ \backup{$\alpha_j^{h'},a$}\;
       	      	}
       	      	$\Gamma_{a,o}^h\,:=\,\textrm{\large $\boxplus$} \Gamma_{a,o_i}^h$\;
       	     }
           $\Gamma_a^{h} \,:=\,R_a \oplus \gamma \cdot \Gamma_{a,o}^h$\;
            $\Gamma^{h} \,:=\, \Gamma^{h} \cup \Gamma_a^{h}$\;
        }  
        	%monahan's pruning first generates all vectors, then prunes
              $V^{h} \,:=\, $\prune{$\Gamma^h$} \;
              $\pi^{*,h} \,:=\, \argmax_{a} \, \Gamma_a^{h}$\;

@@

Line 10 will substitute prime values for all state variables. Line 12 produces the backup $\alpha$-vector and multiplies it by the weight derived from the relevant observation partition. For each observation, two vectors are formed for each corresponding $\alpha$-vector. For example line 12 for observations $O_2,O_3$ and $\alpha_1^1,\alpha_2^1$ is defined as:

\begin{align*}
%\begin{cases}
%\Gamma_{close,O_1,1}^2= 
%\begin{cases}
%(200<t_o<300) \wedge (t_s'>235) &: -900 \\
%(200<t_o<300)\wedge (200<t_s'<235) &: 90 \\
%(200<t_o<300) \wedge (t_s'<200) &: -10 \\
%\end{cases}
%\\
%\Gamma_{close,O_1,2}^2= 
%\begin{cases}
%(200<t_o<300) \wedge (t_s'>200) &: -0.9 \\
%(200<t_o<300) \wedge (t_s'<200) &: -0.1 \\
%\end{cases}
%\end{cases}
%\\
\begin{cases}
\Gamma_{close,O_2,1}^2= 
\begin{cases}
(t_o>300) \wedge (t_s'>235) &: -100 \\
(t_o>300)\wedge (200<t_s'<235) &: 10 \\
(t_o>300) \wedge (t_s'<200) &: 0 \\
\end{cases}
\\
\Gamma_{close,O_2,2}^2= 
\begin{cases}
(t_o>300)\wedge (t_s'>200) &: -0.1 \\
(t_o>300) \wedge (t_s'<200) &: 0 \\
\end{cases}
\end{cases}
\\
\begin{cases}
\Gamma_{close,O_3,1}^2= 
\begin{cases}
(t_o<200) \wedge (t_s'>235) &: -900 \\
(t_o<200)\wedge (200<t_s'<235) &: 90 \\
(t_o<200) \wedge (t_s'<200) &: 0 \\
\end{cases}
\\
\Gamma_{close,O_3,2}^2= 
\begin{cases}
(t_o<200) \wedge (t_s'>200) &: -0.9 \\
(t_o<200)\wedge (t_s'<200) &: 0 \\
\end{cases}
\end{cases}
\end{align*}

Next in line 13 we have to take the cross-sum of the two sets which takes every possible combination of observations according to the $\alpha$-vectors. This means for $O_2 \textrm{\large $\boxplus$} O_3$ we have the following vectors:

\begin{align*}
\begin{cases}
\begin{cases}
(t_s'>235) &: -1000 \\
(200<t_s'<235) &: 100 \\
(t_s'<200) &: 0 \\
\end{cases}
\\
\begin{cases}
(t_s'>235) &: -100.9 \\
(<200<t_s'<235) &: 9.1 \\
(t_s'<200) &: 0 \\
\end{cases}
\\
\begin{cases}
 (t_s'>235) &: -900 \\
 (200<t_s'<235) &: 89.9 \\
 (t_s'<200) &: 0 \\
\end{cases}
\\
\begin{cases}
(t_s'>200) &: -1 \\
(t_s'<200) &: 0 \\
\end{cases}
\end{cases}
\end{align*}

Where the observation partitions ($t_o>300 , t_o<200$) are inconsistent and therefore omitted from the result. These 4 vectors are added with the two vectors from $\Gamma_{close,O_1}^2$ to get $2^3$ different vectors. All 8 vectors are multiplied by the discount $\gamma$ and added with the reward in line 14 where the addition is an augmented $\oplus$ for adding each vector with the reward function. 
% need to add this?
The same operation is performed for the $open$ action and 8 different vectors are formed here too. In line 15 the $\Gamma$ set for each action are added to the set of $\Gamma^h$ to build all the vectors (16 in total) for the next horizon.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\incmargin{.5em}
\linesnumbered
\begin{algorithm}[t!]
\vspace{-.5mm}
\dontprintsemicolon
\SetKwFunction{backup}{Backup}
\SetKwFunction{genObs}{GenRelObs}
\SetKwFunction{prune}{Prune}
\SetKwFunction{remapWithPrimes}{Prime}
\Begin
{
   $V^0:=0, h:=0$\;
   \While{$h < H$}
   {
       $h:=h+1, \Gamma^h :=\emptyset$\;
       \ForEach {$a \in A$}
       {
			$\Gamma_{a}^h :=\emptyset$ \;       		
       		$\textsc{O}^{h-1} \,:=\,$ \genObs{$\Gamma^{h-1},a$}\;
       		 \ForEach {$o_i \in \textsc{O}^{h-1}$}
       		 {
				\ForEach {$\alpha_j^{h-1} \in \Gamma^{h-1}$}
       			{
   	 		  		$\alpha_j^{h'}=$ \remapWithPrimes{$\alpha_j^{h}$} \;
   	 		  		\emph{// $\forall$ $d_{s_i} \to d'_{s_i}$ and $\forall x_{s_i} \to x'_{s_i}$} \; 
   	 		    	$\Gamma_{a,o_i,j}^h \,:=\, P(o_i|xd'_s) \otimes$ \backup{$\alpha_j^{h'},a$}\;
       	      	}
       	      	$\Gamma_{a,o}^h\,:=\,\textrm{\large $\boxplus$} \Gamma_{a,o_i}^h$\;
       	     }
           $\Gamma_a^{h} \,:=\,R_a \oplus \gamma \cdot \Gamma_{a,o}^h$\;
            $\Gamma^{h} \,:=\, \Gamma^{h} \cup \Gamma_a^{h}$\;
        }  
        	%monahan's pruning first generates all vectors, then prunes
              $V^{h} \,:=\, $\prune{$\Gamma^h$} \;
              $\pi^{*,h} \,:=\, \argmax_{a} \, \Gamma_a^{h}$\;

@@

Line 10 will substitute prime values for all state variables. Line 12 produces the backup $\alpha$-vector and multiplies it by the weight derived from the relevant observation partition. For each observation, two vectors are formed for each corresponding $\alpha$-vector. For example line 12 for observations $O_1,O_2,O_3$ and $\alpha_1^1,\alpha_2^1$ is defined as:

\begin{align*}
\begin{cases}
\Gamma_{close,O_1,1}^2= 
\begin{cases}
(t_s'>235) &: -900 \\
(200<t_s'<235) &: 90 \\
(t_s'<200) &: -10 \\
\end{cases}
\\
\Gamma_{close,O_1,2}^2= 
\begin{cases}
(t_s'>200) &: -0.9 \\ 
(t_s'<200) &: -0.1 \\
\end{cases}
\end{cases}
\\
\begin{cases}
\Gamma_{close,O_2,1}^2= 
\begin{cases}
 (t_s'>235) &: -100 \\
(200<t_s'<235) &: 10 \\
(t_s'<200) &: 0 \\
\end{cases}
\\
\Gamma_{close,O_2,2}^2= 
\begin{cases}
 (t_s'>200) &: -0.1 \\
 (t_s'<200) &: 0 \\
\end{cases}
\end{cases}
\\
\begin{cases}
\Gamma_{close,O_3,1}^2= 
\begin{cases}
 (t_s'>235) &: -900 \\
 (200<t_s'<235) &: 90 \\
 (t_s'<200) &: 0 \\
\end{cases}
\\
\Gamma_{close,O_3,2}^2= 
\begin{cases}
 (t_s'>200) &: -0.9 \\
(t_s'<200) &: 0 \\
\end{cases}
\end{cases}
\end{align*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.4\textwidth]{policy2.pdf}
\end{center}
\vspace{-3mm}
\caption{%\footnotesize 
An example of a policy tree in \textsc{Power Plant}
for the $close$ action, $O_1,O_3$ based on $\alpha_1^1$ and $O_2$ based on $\alpha_2^1$.} 
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Next in line 13 we have to take the cross-sum of the two sets which takes every possible combination of observations according to the $\alpha$-vectors. This means for $O_1 \textrm{\large $\boxplus$} O_2 \textrm{\large $\boxplus$} O_3$ we have 8 vectors where the first 4 combination  have been shown below:

\begin{align*}
\begin{cases}
\begin{cases}
(t_s'>235) &: -1900 \\
(200<t_s'<235) &: 190 \\
(t_s'<200) &: -10 \\
\end{cases}
\\
\begin{cases}
(t_s'>235) &: -1000.9 \\
(<200<t_s'<235) &: 99.1 \\
(t_s'<200) &: -10 \\
\end{cases}
\\
\begin{cases}
 (t_s'>235) &: -1800.1 \\
 (200<t_s'<235) &: 179.9 \\
 (t_s'<200) &: -10 \\
\end{cases}
\\
\begin{cases}
 (t_s'>235) &: -901 \\
 (200<t_s'<235) &: 89 \\
 (t_s'<200) &: -10 \\
\end{cases}
\end{cases}
\end{align*}

We have shown the policy tree for the third vector in figure 3. The weights here are distributed according to each of the observations while the first and last partition choose to take a backup from the $\alpha_1^1$ and the middle partition uses $\alpha_2^1$. 
All 8 vectors are multiplied by the discount $\gamma$ and added with the reward in line 14 where the addition is an augmented $\oplus$ for adding each vector with the reward function.

\begin{align*}
\begin{cases}
\begin{cases}
(t_s'>235) &: -2900 \\
(200<t_s'<235) &: -810 \\
(t_s'<200) &: 90 \\
\end{cases}
\\
\begin{cases}
(t_s'>235) &: -2000.9 \\
(<200<t_s'<235) &: -900.9 \\
(t_s'<200) &: 90 \\
\end{cases}
\\
\begin{cases}
 (t_s'>235) &: -2800.1 \\
 (200<t_s'<235) &: -820.1 \\
 (t_s'<200) &: 90 \\
\end{cases}
\\
\begin{cases}
 (t_s'>235) &: -1901 \\
 (200<t_s'<235) &: -911 \\
 (t_s'<200) &: 90 \\
\end{cases}
\end{cases}
\end{align*}

The same operation is performed for the $open$ action and 8 different vectors are formed here too. In line 15 the $\Gamma$ set for each action are added to the set of $\Gamma^h$ to build all the vectors (16 in total) for the next horizon.
